Appendix: New and simplified manual controls for projection and slice tours

### 2022-10-04 paper/appendix.Rmd
Fixing the appendix

## Adjustment method 3

Finally we may also consider an adjustment that works within the projection plane. The idea is that when applying Gram-Schmidt, the direction of the first basis vector is not changed. We thus rotate the projection matrix within the plane such that the first basis is aligned with the direction of the selected variable $m$ before applying orthonormalization. In addition we need to keep the contribution of variable $m$ along that direction fixed when normalizing the vector. The following steps are needed:

1. Provide $A$, and $m$.
2. Change values in row $m$, giving $A^*$.
3. Rotate the basis within the projection plane, such that the first basis direction is aligned with the direction of $m$ in $A^*$, giving $\hat A^*$. In a 2D projection this means finding the angle $\theta$ by which the basis should be rotated.
4. Normalize the first direction in $\hat A^*$ while keeping the contribution in row $m$ fixed, giving $\hat A^{*0}$.
5. Orthonormalize $\hat A^{*0}$ using Gram-Schmidt, giving $\hat A^{**}$. By construction the contribution of $m$ will remain fixed.
6. Rotate $\hat A^{**}$ back into the original basis to obtain $A^{**}$.

Diagram needed?

{\textcolor{blue} Ursula, I am more convinced now that this is exactly method 2, the only difference is that you want to rotate to get to the "first" direction, whereas I simply relabel the rows of the matrix}

@@

The slice center, $c$, from Equation 1 in the main paper, is a point in the $p$-dimensional space. By default it is placed in the center of the data (Figure \ref{fig:anchornav} A). Moving this along a variable axis moves the slice outwards from the center of the data (Figure \ref{fig:anchornav} B, C).

@@

The Slice Tour* notebooks include the additional functions: `ProjectionPlot`, `Projected2DSliderPlot`, and `VisualiseSliceDyanmic`.

\textcolor{red}{does this mean that we are changing the name of the package from mmtour.wl to Slice Tour notebook?}

>>>>

\textcolor{red}{We could also take a similar approach without completing the basis, through rotation within the $d$-dimensional hyperplane defined by the new projection. In that case we would again start from $A^*$ as above, and then rotate the basis such that the first direction is aligned with the selected variable $m$. Applying Gram-Schmid within that basis definition will not alter the direction for the first vector, thus the direction of $m$ remains fixed to what was selected with the cursor. By adjusting the normalization procedure we can also ensure that the lenght remains as selcted manually by the user. The advantage is that we do not introduce random jumps from the other variables, but keep changes minimal within the $p\times d$ projection matrix.}

@@

The slice center, $c$, from Equation 1 in the main paper, is a point in the $p$-dimensional space. By default, it is placed in the center of the data (Figure \ref{fig:anchornav} A). Moving this along a variable axis moves the slice outwards from the center of the data (Figure \ref{fig:anchornav} B, C). The visual display of the center position is a form of star plot as suggested useful in @condviz2, where the position of the point is marked by a polygon on radial axes indicating its position relative to the variable minimum to maximum values.

@@

The package includes the additional functions not used in the example application: `ProjectionPlot`, `Projected2DSliderPlot`, and `VisualiseSliceDyanmic`.

### 2022-10-04 paper/appendix.tex
Fixing the appendix

\hypertarget{adjustment-method-3}{%
\subsection{Adjustment method 3}\label{adjustment-method-3}}

Finally we may also consider an adjustment that works within the
projection plane. The idea is that when applying Gram-Schmidt, the
direction of the first basis vector is not changed. We thus rotate the
projection matrix within the plane such that the first basis is aligned
with the direction of the selected variable \(m\) before applying
orthonormalization. In addition we need to keep the contribution of
variable \(m\) along that direction fixed when normalizing the vector.
The following steps are needed:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Provide \(A\), and \(m\).
\item
  Change values in row \(m\), giving \(A^*\).
\item
  Rotate the basis within the projection plane, such that the first
  basis direction is aligned with the direction of \(m\) in \(A^*\),
  giving \(\hat A^*\). In a 2D projection this means finding the angle
  \(\theta\) by which the basis should be rotated.
\item
  Normalize the first direction in \(\hat A^*\) while keeping the
  contribution in row \(m\) fixed, giving \(\hat A^{*0}\).
\item
  Orthonormalize \(\hat A^{*0}\) using Gram-Schmidt, giving
  \(\hat A^{**}\). By construction the contribution of \(m\) will remain
  fixed.
\item
  Rotate \(\hat A^{**}\) back into the original basis to obtain
  \(A^{**}\).
\end{enumerate}

Diagram needed?

\{\textcolor{blue} Ursula, I am more convinced now that this is exactly
method 2, the only difference is that you want to rotate to get to the
``first'' direction, whereas I simply relabel the rows of the matrix\}

@@

The slice center, \(c\), from Equation 1 in the main paper, is a point
in the \(p\)-dimensional space. By default it is placed in the center of
the data (Figure \ref{fig:anchornav} A). Moving this along a variable
axis moves the slice outwards from the center of the data (Figure
\ref{fig:anchornav} B, C).

@@

The Slice Tour* notebooks include the additional functions:
\texttt{ProjectionPlot}, \texttt{Projected2DSliderPlot}, and
\texttt{VisualiseSliceDyanmic}.

\textcolor{red}{does this mean that we are changing the name of the package from mmtour.wl to Slice Tour notebook?}

>>>>

\textcolor{red}{We could also take a similar approach without completing the basis, through rotation within the $d$-dimensional hyperplane defined by the new projection. In that case we would again start from $A^*$ as above, and then rotate the basis such that the first direction is aligned with the selected variable $m$. Applying Gram-Schmid within that basis definition will not alter the direction for the first vector, thus the direction of $m$ remains fixed to what was selected with the cursor. By adjusting the normalization procedure we can also ensure that the lenght remains as selcted manually by the user. The advantage is that we do not introduce random jumps from the other variables, but keep changes minimal within the $p\times d$ projection matrix.}

@@

The slice center, \(c\), from Equation 1 in the main paper, is a point
in the \(p\)-dimensional space. By default, it is placed in the center
of the data (Figure \ref{fig:anchornav} A). Moving this along a variable
axis moves the slice outwards from the center of the data (Figure
\ref{fig:anchornav} B, C). The visual display of the center position is
a form of star plot as suggested useful in \citet{condviz2}, where the
position of the point is marked by a polygon on radial axes indicating
its position relative to the variable minimum to maximum values.

@@

The package includes the additional functions not used in the example
application: \texttt{ProjectionPlot}, \texttt{Projected2DSliderPlot},
and \texttt{VisualiseSliceDyanmic}.
